---
title: "[Your title (e.g Exercise 4.3.2)]"
author:
  - name: [First author name]
    #url: https://example.com/norajones
    affiliation: cand.merc (OSCM)
    #affiliation_url: https://example.com/spacelysprokets
  - name: [Second author name]
    affiliation: cand.merc (?)
  - name: [Third author name]
    affiliation: cand.merc (?)
output: 
  html_document:
    toc: true
    toc_depth: 2
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  autodep = TRUE
  )
```

<!-- Some css for the comments (delete) -->
```{css, echo=FALSE}
.comment {
  color: gold;
  border-style: dashed;
  border-width: thin;
  padding: 10px;
}
```


<div class="comment">This is a template for a TFA exercise. Modify it as needed. Text in gold are comments. Some example text and code is given. Delete as needed.</div>

## Introduction

We load the following packages:

```{r}
library(tidyverse)
```

<div class="comment">
Maybe a short intro about the exercise and how data are imported.

Some help links:

- To use this template you must have installed the tfa package (`remotes::install_github("bss-osca/tfa/tfa-package")`). Go to **File > New File > R Markdown...** and select the **From template** in the left column and then **TFA Exercise**.
- For Markdown basics see **Help > Markdown Quick Reference** in RStudio.
- For R Markdown basics see the **Help > Cheatsheets > R Markdown Cheat Sheet** in RStudio.
- Learn more about he Distill format for R Markdown at <https://rstudio.github.io/distill>.
- To see the possible options for R chunks see <https://yihui.org/knitr/options/>.
</div>

Consider a problem of determining the best sequencing of jobs on a machine. A set of startup costs are given for 5 machines:

```{r}
startup_costs <- c(27, 28, 32, 35, 26)
startup_costs
```

Moreover, when changing from one job to another job the setup costs are given as:

```{r}
setup_costs <- matrix(c(
  NA, 35, 22, 44, 12,
  49, NA, 46, 38, 17,
  46, 12, NA, 29, 41,
  23, 37, 31, NA, 26,
  17, 23, 28, 34, NA), 
  byrow = T, nrow = 5)
setup_costs
```

The goal of the problem is to determine a sequence of jobs which minimizes the total setup cost including the startup cost.

One possible way to find a sequence is the use a greedy strategy:

```
Greedy Algorithm
Step 0: Start with the job which has minimal startup cost.
Step 1: Select the next job as the job not already done 
        with minimal setup cost given current job. 
Step 2: Set next job in Step 1 to current job and 
        go to Step 1 if not all jobs are done.
```

The greedy algorithm can be implemented as:

```{r}
greedy <- function(startup, setup) {
  jobs <- nrow(setup)
  cur_job <- which.min(startup)
  cost <- startup[cur_job]
  cat("Start job:", cur_job, "\n")
  job_seq <- cur_job
  setup[, cur_job] <- NA
  # browser()
  for (i in 1:(jobs-1)) {
    next_job <- which.min(setup[cur_job, ])
    cat("Next job:", next_job, "\n") 
    cost <- cost + setup[cur_job, next_job]
    job_seq <- c(job_seq, next_job)
    cur_job <- next_job
    setup[, cur_job] <- NA
  }
  # print(setup)
  return(list(seq = job_seq, cost = cost))
}
greedy(startup_costs, setup_costs)
```

First, the job with minimum startup cost is found using function `which.min` and we define cost a the startup cost. We use `cat` to make some debugging statements and initialize `job_seq` with the first job. Next we have to have a way ignoring jobs already done. We do that here by setting the columns of setup cost equal to NA for jobs already done. Hence they will not be selected by `which.min`. The `for` loop runs 4 times and selects jobs and accumulate the total cost.


## Question 1

A well-known better strategy is to:

```
Better Algorithm
Step 0: Subtract minimum of startup and setup cost for each job from setup and 
        startup costs (that is columnwise)
Step 1: Call the greedy algorithm with the modified costs. Note that the total 
        cost returned has to be modified a bit.
```

*Implement a `better` function calculating a better strategy. Hint: to find the minimum column costs you may use `apply(rbind(startup, setup), 2, min, na.rm = T)`.*

The function becomes:

```{r}
better <- function(startup, setup) {
  jobs <- nrow(setup)
  min_col_val <- apply(rbind(startup, setup), 2, min, na.rm = T)
  startup <- startup - min_col_val
  min_mat <- matrix(rep(min_col_val, jobs), 
                    ncol = jobs, byrow = T)
  setup <- setup - min_mat
  lst <- greedy(startup, setup)
  lst$cost <- lst$cost + sum(min_col_val)
  return(lst)
}
```

If we run the function the job sequence becomes:

```{r}
res <- better(startup_costs, setup_costs)
```

That is we use job sequence `r res$seq` with a cost of `r res$cost`.


## Question 2

*Assume that the startup costs are 27, 28, 55, 120 and 26 instead. How will that affect the job sequence using the better strategy?*

We modify the startup costs and run `better`:

```{r}
startup_costs <- c(27, 28, 55, 120, 26)
res <- better(startup_costs, setup_costs)
```

The job sequence now becomes `r res$seq` with a cost of `r res$cost`.


## Question 3

*Make a plot of the job sequence*

```{r, layout="l-body-outset", fig.asp = 0.25, eval=FALSE}
library(ggraph)
library(tidygraph)
edges <- tibble(from = res$seq[1:4], to = res$seq[2:5])
graph <- as_tbl_graph(edges) 
ggraph(graph, layout = 'linear') + 
  geom_edge_link(arrow = arrow(length = unit(4, 'mm')), 
                 end_cap = circle(4, 'mm'), 
                 start_cap  = circle(4, 'mm')) + 
  geom_node_point(colour = "gray", size = 9, alpha = 0.5) +
  geom_node_text(aes(label=name)) +
  theme_graph()
```


## Colophon 
<!-- Always keep this section for reproducibility -->

This report has been created inside [RStudio](http://www.rstudio.com/ide/) using [R Markdown](https://rmarkdown.rstudio.com/). 

The report was built using:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session <- devtools::session_info()
session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages %>% 
  select(package, loadedversion, date, source) %>% 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = 5, 
                           autoWidth = FALSE,
                           bInfo = FALSE,
                           paging = TRUE))
```

```{r}
library(gapminder)
library(tidyverse)
gapminder%>%
  glimpse()%>%
  summary()

gapminder%>%
  count(country)%>%
  nrow()
gapminder%>%
  count(continent)%>%
  nrow()
gapminder%>%
  count(continent,country)%>%
  count(continent)

gapminder%>%
  distinct(continent)%>%
  pull(continent)%>%
  as.character()

gapminder%>%
  filter(lifeExp<29)
gapminder%>%
  filter(country=="Rwanda",  year>1979)
gapminder%>%
  filter(country %in% c("Rwanda","Afghanistan","France"))

gapminder%>%
  select(year,lifeExp,country,gdpPercap)

gapminder%>%
  filter(continent=="Africa"& gdpPercap<500|continent=="Europe" & gdpPercap>40000)%>%
  select(continent,gdpPercap,country)

gapminder%>%
  mutate(country_gdp=pop*gdpPercap)

mean_dk <- gapminder %>% 
  filter(country == "Denmark") %>% 
  pull(gdpPercap) %>% 
  mean()
dat <- gapminder %>% 
  mutate(gdpPercapRel = gdpPercap/mean_dk)


gapminder%>%
  arrange(year,country)
gapminder%>%
  filter(year=="2007")%>%
  arrange(desc(lifeExp))

gapminder%>%
  select(everything(),yr=year,-pop)
gapminder%>%
  select(year,pop,everything())

gapminder%>%
  group_by(continent)%>%
 summarise(n=n())

gapminder%>%
  group_by(continent)%>%
 summarise(n=n(),con=n_distinct(country),mean_life=mean(lifeExp))  
gapminder%>%
  filter(continent=="Asia")%>%
  group_by(year)%>%
  summarize(Minn=min(lifeExp),max=max(lifeExp))


gapminder %>%
  group_by(country)%>%
  select(country,year,lifeExp)%>%
  arrange(year,.by_group=TRUE)%>%
  mutate(gain_life = lifeExp - first(lifeExp))%>%
  filter(year<1963)

gapminder %>%
  select(country, year, continent, lifeExp) %>%             #
  group_by(continent, country) %>% 
  mutate(le_delta = lifeExp - lag(lifeExp)) %>%  # lag tager perioden for saa det er nuværende periode minus forhenvaerende periode         
  summarize(worst_le_delta = min(le_delta, na.rm = TRUE)) %>%    # 
  slice_min(worst_le_delta) %>%                             # 
  arrange(worst_le_delta)

install.packages("babynames")
library(babynames)
library(skimr)
skim(babynames)

babynames%>%
  select(name,prop)%>%
  filter(prop>=0.08)
babynames%>%
  filter(name=="Sea")

babynames%>%
  filter(name=="Sue",sex=="M")

babynames%>%
  filter(year=="1880",n=="5"|n=="6")
babynames%>%
  filter(name%in%c("Acura","Lexus","Yugo"))

min(babynames$n)

babynames%>%
  filter(year=="2017",sex=="F")%>%
  select(name,n)%>%
  arrange(desc(n))

babynames%>%
  filter(name=="Khaleesi")%>%
  summarise(sum(n),year=first(year))

babynames%>%
  group_by(name,sex)%>%
  summarise(pop=sum(n))%>%
  arrange(desc(pop))

babynames%>%
  group_by(year)%>%
  summarise(sum(n))



babynames%>%
  group_by(year,sex)%>%
  mutate(rank=min_rank(desc(prop)))%>%# prop kan skiftes med n men giver samme resultat
  arrange(year,sex,desc(prop))%>%
filter(rank=="1",year>2009)

library(skimr)
library(tidyverse)
path <- system.file("extdata/profit_raw.csv", package = "tfa")
profit <- read_csv(path)
skim(profit)

profit <- profit %>%  
  mutate(Quarter = str_remove(Quarter, "q") %>% 
           str_remove("Q") %>% 
           as.numeric())
profit %>% distinct(Quarter)

profit <- profit %>% 
  mutate(Revenue = str_replace_all(Revenue, c("unknown" = "?", "Unknown" = "?"))) %>% 
  mutate(Revenue = na_if(Revenue, "?")) %>% 
  mutate(Revenue = str_remove(Revenue, fixed("$ ")) %>% as.numeric())

profit%>%
  mutate(`Customer Service Cost`=str_remove(`Customer Service Cost`,fixed("$ "))%>% as.numeric())%>%
  mutate(Profit=str_remove(Profit,fixed("$ "))%>% as.numeric())%>%
  mutate(`Product Cost`=str_remove(`Product Cost`,fixed("$ "))%>% as.numeric())

profit <- read_csv(path) %>% 
  mutate(Quarter = str_remove(Quarter, "q") %>% str_remove("Q") %>% as.numeric()) %>% 
  mutate(across(Revenue:Profit, ~str_replace_all(.x, c("unknown" = "?", "Unknown" = "?")) )) %>% 
  mutate(across(Revenue:Profit, ~na_if(.x, "?") )) %>% 
  mutate(across(Revenue:Profit, ~str_remove(.x, fixed("$ ")) %>% as.numeric() )) 
profit
skim(profit)

profit <- profit %>% 
  mutate(Profit_calc = Revenue - `Product Cost` - `Customer Service Cost`, Diff = abs(Profit - Profit_calc)) %>%  
  mutate(Profit = if_else(Diff > 0 & Diff <= 1, Profit_calc, Profit, missing = Profit) ) 
profit %>% filter(Diff == 1, is.na(Profit_calc))

profit <- profit %>% 
  rowwise() %>% 
  mutate(c_na = sum(is.na(c_across(Revenue:Profit)))) %>% 
  mutate(Revenue = if_else(is.na(Revenue) & c_na == 1, Profit + `Product Cost` + `Customer Service Cost`, Revenue, Revenue),
         `Product Cost` = if_else(is.na(`Product Cost`) & c_na == 1, - Profit + Revenue - `Customer Service Cost`, `Product Cost`), 
         `Customer Service Cost` = if_else(is.na(`Customer Service Cost`) & c_na == 1, - Profit + Revenue - `Product Cost`, `Customer Service Cost`)) %>% 
  select(Quarter:Profit)

profit %>% 
  mutate(Profit_calc = Revenue - `Product Cost` - `Customer Service Cost`, Diff = abs(Profit - Profit_calc)) %>%  
  filter(Diff > 0)
 
profit %>% 
  rowwise() %>% #perform on each row at a time. good when no vector function is avaiable and is used with c_across
  mutate(c_na = sum(is.na(c_across(Revenue:Profit)))) %>% 
  filter(c_na > 0) 

profit%>%
  select(Profit,Quarter)%>%
  group_by(Quarter)%>%
  slice_max(Profit,n=2)
  
profit%>%
  group_by(Quarter,`Customer ID`)%>%
  summarise(Profit=sum(Profit))%>%
  slice_max(Profit,n=2)

profit%>%
  group_by(`Product Line`)%>%
  summarise(Profit=sum(Profit))%>%
  slice_max(Profit)
profit%>%
  group_by(Channel)%>%
  summarise(Profit=sum(Profit))%>%
  slice_max(Profit)  

#orrrr
library(purrr)
val <- names(profit)[1:5]
max_profit <- map_df(
  val, 
  ~{
      tmp <- profit %>% 
        group_by(.data[[.x]]) %>% 
        summarise(Profit = sum(Profit), .groups = "drop") %>% 
        slice_max(Profit)
      tibble(by = .x, 
        best = as.character(tmp[[1,1]]),
        profit = tmp[[1,2]] 
      )
    } 
  )
max_profit 

profit%>%
  group_by(`Customer ID`)%>%
  distinct(Country)%>%
  count(`Customer ID`)

profit%>%
  arrange(desc(Profit),desc(Revenue))

profit%>%
  group_by(`Product Line`)%>%
  summarise(Cost=sum(`Product Cost`+`Customer Service Cost`))%>%
  slice_max(Cost)
  
profit%>%
  group_by(`Product Line`)%>%
  summarise(Cost=sum(`Product Cost`))%>%
  slice_min(Cost)

profit %>%
  mutate(cust_cost_new = `Customer Service Cost` * 1.05, 
         profit_new = Revenue - cust_cost_new - `Product Cost`) %>% 
  group_by(`Product Line`) %>% 
  summarise(cust_cost = sum(`Customer Service Cost`), profit = sum(Profit), 
            cust_cost_new = sum(cust_cost_new), profit_new = sum(profit_new),
            profit_decrease = profit_new - profit)
  

```








